---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# GenomicState

[![Travis-CI build status](https://travis-ci.org/LieberInstitute/GenomicState.svg?branch=master)](https://travis-ci.org/LieberInstitute/GenomicState)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/LieberInstitute/GenomicState?branch=master&svg=true)](https://ci.appveyor.com/project/LieberInstitute/GenomicState)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Codecov test coverage](https://codecov.io/gh/LieberInstitute/GenomicState/branch/master/graphs/badge.svg)](https://codecov.io/gh/LieberInstitute/GenomicState?branch=master)

This package provides functions for building `GenomicState` objects from diverse annotation sources such as [`Gencode`](https://www.gencodegenes.org/human/releases.html). It also provides a way to load pre-computed `GenomicState` objects if you are working at [JHPCE](https://jhpce.jhu.edu/). These `GenomicState` objects are normally created using [derfinder::makeGenomicState()](https://rdrr.io/bioc/derfinder/man/makeGenomicState.html) and can be used for annotating regions with [derfinder::annotateRegions()](https://rdrr.io/bioc/derfinder/man/annotateRegions.html) which are in turn used by [derfinderPlot::plotRegionCoverage()](https://rdrr.io/bioc/derfinderPlot/man/plotRegionCoverage.html).

# Installation instructions

Get the latest stable `R` release from [CRAN](http://cran.r-project.org/). Then install `GenomicState` using the following code:

```{r, eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("GenomicState")
```


# Citation

Below is the citation output from using `citation('GenomicState')` in R. Please 
run this yourself to check for any updates on how to cite __GenomicState__.

```{r 'citation'}
citation('GenomicState')
```

# Testing

Testing on Bioc-devel is feasible thanks to [R Travis](http://docs.travis-ci.com/user/languages/r/).
