---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/"
)
```

# GenomicState

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![BioC status](http://bioconductor.org/images/shields/availability/all.svg)](http://bioconductor.org/packages/release/data/annotation/html/GenomicState.html#archives)
[![Codecov test coverage](https://codecov.io/gh/LieberInstitute/GenomicState/branch/master/graph/badge.svg)](https://codecov.io/gh/LieberInstitute/GenomicState?branch=master)
[![R build status](https://github.com/LieberInstitute/GenomicState/workflows/R-CMD-check-bioc/badge.svg)](https://github.com/LieberInstitute/GenomicState/actions)
[![Support site activity, last 6 months: tagged questions/avg. answers per question/avg. comments per question/accepted answers, or 0 if no tagged posts.](http://www.bioconductor.org/shields/posts/GenomicState.svg)](https://support.bioconductor.org/t/GenomicState/)
[![GitHub issues](https://img.shields.io/github/issues/LieberInstitute/GenomicState)](https://github.com/LieberInstitute/GenomicState/issues)
<!-- badges: end -->

This package provides functions for building `GenomicState` objects from diverse annotation sources such as [`Gencode`](https://www.gencodegenes.org/human/releases.html). It also provides a way to load pre-computed `GenomicState` objects if you are working at [JHPCE](https://jhpce.jhu.edu/). These `GenomicState` objects are normally created using [derfinder::makeGenomicState()](http://lcolladotor.github.io/derfinder/reference/makeGenomicState.html) and can be used for annotating regions with [derfinder::annotateRegions()](http://lcolladotor.github.io/derfinder/reference/annotateRegions.html) which are in turn used by [derfinderPlot::plotRegionCoverage()](http://leekgroup.github.io/derfinderPlot/reference/plotRegionCoverage.html).

## Documentation

For more information about `GenomicState` check the vignettes [through Bioconductor](http://bioconductor.org/packages/GenomicState) or at the [documentation website](http://lieberinstitute.github.io/GenomicState).

## Installation instructions

Get the latest stable `R` release from [CRAN](http://cran.r-project.org/). Then install `GenomicState` from [Bioconductor](http://bioconductor.org/) using the following code:

```{r 'install', eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
      install.packages("BiocManager")
  }

BiocManager::install("GenomicState")
```

## Citation

Below is the citation output from using `citation('GenomicState')` in R. Please 
run this yourself to check for any updates on how to cite __GenomicState__.

```{r 'citation', eval = requireNamespace('GenomicState')}
print(citation("GenomicState"), bibtex = TRUE)
```

Please note that the `GenomicState` was only made possible thanks to many other R and bioinformatics software authors, which are cited either in the vignettes and/or the paper(s) describing this package.

## Code of Conduct
  
Please note that the derfinderPlot project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

## Development tools

* Continuous code testing is possible thanks to [GitHub actions](https://www.tidyverse.org/blog/2020/04/usethis-1-6-0/)  through `r BiocStyle::CRANpkg('usethis')`, `r BiocStyle::CRANpkg('remotes')`, `r BiocStyle::Githubpkg('r-hub/sysreqs')` and `r BiocStyle::CRANpkg('rcmdcheck')` customized to use [Bioconductor's docker containers](https://www.bioconductor.org/help/docker/) and `r BiocStyle::Biocpkg('BiocCheck')`.
* Code coverage assessment is possible thanks to [codecov](https://codecov.io/gh) and `r BiocStyle::CRANpkg('covr')`.
* The [documentation website](http://lieberinstitute.github.io/GenomicState) is automatically updated thanks to `r BiocStyle::CRANpkg('pkgdown')`.
* The code is styled automatically thanks to `r BiocStyle::CRANpkg('styler')`.
* The documentation is formatted thanks to `r BiocStyle::CRANpkg('devtools')` and `r BiocStyle::CRANpkg('roxygen2')`.

For more details, check the `dev` directory.

This package was developed using `r BiocStyle::Biocpkg('biocthis')`.
