---
title: "Introduction to GenomicState"
author: 
  - name: Leonardo Collado-Torres
    affiliation:
    - &libd Lieber Institute for Brain Development, Johns Hopkins Medical Campus
    - &ccb Center for Computational Biology, Johns Hopkins University
    email: lcolladotor@gmail.com
date: "`r doc_date()`"
package: "`r pkg_ver('GenomicState')`"
output: 
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Introduction to GenomicState}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

```{r vignetteSetup, echo=FALSE, message=FALSE, warning = FALSE}
## For links
library('BiocStyle')

## Track time spent on making the vignette
startTime <- Sys.time()

## Bib setup
library('knitcitations')

## Load knitcitations with a clean bibliography
cleanbib()
cite_options(hyperlink = 'to.doc', citation_format = 'text', style = 'html')
# Note links won't show for now due to the following issue
# https://github.com/cboettig/knitcitations/issues/63

## Write bibliography information
bib <- c(
    R = citation(),
    BiocStyle = citation('BiocStyle'),
    GenomeInfoDb = RefManageR::BibEntry(bibtype = 'manual',
        key = 'GenomeInfoDb',
        author = 'Sonali Arora and Martin Morgan and Marc Carlson and H. Pagès',
        title = "GenomeInfoDb: Utilities for manipulating chromosome and other 'seqname' identifiers",
        year = 2017, doi = '10.18129/B9.bioc.GenomeInfoDb'),
    knitcitations = citation('knitcitations'),
    knitr = citation('knitr')[3],
    GenomicState = citation('GenomicState')[1],
    rmarkdown = citation('rmarkdown'),
    rtracklayer = citation('rtracklayer')[1],
    sessioninfo = citation('sessioninfo'),
    testthat = citation('testthat'),
    GenomicFeatures = citation('GenomicFeatures'),
    bumpunter = citation('bumphunter')[1],
    derfinder = citation('derfinder')[1],
    AnnotationDbi = citation('AnnotationDbi'),
    IRanges = citation('IRanges'),
    org.Hs.eg.db = citation('org.Hs.eg.db'),
    glue = citation('glue'),
    AnnotationHub = citation('AnnotationHub'),
    AnnotationHubData = citation('AnnotationHubData'),
    GenomicRanges = citation('GenomicRanges')
)
```

# Basics

## Install `r Biocpkg('GenomicState')`

`R` is an open-source statistical environment which can be easily modified to enhance its functionality via packages. `r Biocpkg('GenomicState')` is a `R` package available via GitHub. `R` can be installed on any operating system from [CRAN](https://cran.r-project.org/) after which you can install `r Biocpkg('GenomicState')` by using the following commands in your `R` session:

```{r 'install', eval = FALSE}
## If necessary
install.packages('remotes')

## Install GenomicState
remotes::install_github('LieberInstitute/GenomicState')
```

## Required knowledge

`r Biocpkg('GenomicState')` is based on many other packages and in particular in those that have implemented the infrastructure needed for dealing with annotation data. That is, packages like `r Biocpkg('rtracklayer')` that allow you to import the data. A `r Biocpkg('GenomicState')` user is not expected to deal with those packages directly but will need to be familiar with `r Biocpkg('derfinder')` and `r Biocpkg('derfinderPlot')` to understand the results `r Biocpkg('GenomicState')` generates. Furthermore, it'll be useful for the user to know the syntax of `r Biocpkg('AnnotationHub')` `r citep(bib[['AnnotationHub']])` in order to query and load the data provided by this package.

If you are asking yourself the question "Where do I start using Bioconductor?" you might be interested in [this blog post](http://lcolladotor.github.io/2014/10/16/startbioc/#.VkOKbq6rRuU).

## Asking for help

As package developers, we try to explain clearly how to use our packages and in which order to use the functions. But `R` and `Bioconductor` have a steep learning curve so it is critical to learn where to ask for help. The blog post quoted above mentions some but we would like to highlight the [Bioconductor support site](https://support.bioconductor.org/) as the main resource for getting help regarding Bioconductor. Other alternatives are available such as creating GitHub issues and tweeting. However, please note that if you want to receive help you should adhere to the [posting guidelines](http://www.bioconductor.org/help/support/posting-guide/). It is particularly critical that you provide a small reproducible example and your session information so package developers can track down the source of the error.

## Citing `r Biocpkg('GenomicState')`

We hope that `r Biocpkg('GenomicState')` will be useful for your research. Please use the following information to cite the package and the overall approach. Thank you!

```{r 'citation'}
## Citation info
citation('GenomicState')
```

# Overview

The `r Biocpkg('GenomicState')` package was developed for speeding up analyses that require these objects and in particular those that rely on Gencode annotation data. The package `r Biocpkg('GenomicState')` provides functions for building `GenomicState` objects from diverse annotation sources such as [`Gencode`](https://www.gencodegenes.org/human/releases.html). It also provides a way to load pre-computed `GenomicState` objects if you are working at [JHPCE](https://jhpce.jhu.edu/). These `GenomicState` objects are normally created using [derfinder::makeGenomicState()](https://rdrr.io/bioc/derfinder/man/makeGenomicState.html) and can be used for annotating regions with [derfinder::annotateRegions()](https://rdrr.io/bioc/derfinder/man/annotateRegions.html) which are in turn used by [derfinderPlot::plotRegionCoverage()](https://rdrr.io/bioc/derfinderPlot/man/plotRegionCoverage.html).


To get started, load the `r Biocpkg('GenomicState')` package.


```{r setup, message = FALSE, warning = FALSE}
library('GenomicState')
```

# AnnotationHub

This is how you can query the data from `GenomicState` using `r Biocpkg('AnnotationHub')` `r citep(bib[['AnnotationHub']])`.

```{r 'annotation_hub'}
## Query AnnotationHub for the GenomicState object for Gencode v31 on
## hg19 coordinates
hub_gs_gencode_v31_hg19 <- GenomicStateHub(version = '31', genome = 'hg19',
    filetype = 'GenomicState')
hub_gs_gencode_v31_hg19

## Check the metadata
mcols(hub_gs_gencode_v31_hg19)

## Access the file through AnnotationHub
if(length(hub_gs_gencode_v31_hg19) == 1) {
    gs_gencode_v31_hg19 <- hub_gs_gencode_v31_hg19[[1]]
}
```

(This section is currently incomplete!)

# Using the objects

To show how we can use these objects, first we build those for Gencode version 31 on hg19 coordinates.

```{r 'build_ex_objects'}
## Load the example TxDb object
## or start from scratch with:
## txdb_v31_hg19_chr21 <- gencode_txdb(version = '31', genome = 'hg19',
##     chrs = 'chr21')
txdb_v31_hg19_chr21 <- AnnotationDbi::loadDb(
    system.file('extdata', 'txdb_v31_hg19_chr21.sqlite',
        package = 'GenomicState')
)

## Build the GenomicState and annotated genes
genes_v31_hg19_chr21 <- gencode_annotated_genes(txdb_v31_hg19_chr21)
gs_v31_hg19_chr21 <- gencode_genomic_state(txdb_v31_hg19_chr21)
```

Next we load a series of related packages that use these objects.

```{r 'load_derfinder', messages = FALSE, warnings = FALSE}
## Load external packages
library('derfinder')
library('derfinderPlot')
library('bumphunter')
library('GenomicRanges')
```

Next we can prepare the needed for running `derfinderPlot::plotRegionCoverage()` where we use the `TxDb` object, the `GenomicState` and the `annotated genes` we prepared for Gencode v31 on hg19.

```{r 'example_plot_prep'}
## Some example regions from derfinder (set the chromosome lengths)
regions <- genomeRegions$regions[1:2]
seqlengths(regions) <- seqlengths(txdb_v31_hg19_chr21)[
    names(seqlengths(regions))
]

## Annotate them
nearestAnnotation <- matchGenes(x = regions, subject = genes_v31_hg19_chr21)
annotatedRegions <- annotateRegions(regions = regions, 
    genomicState = gs_v31_hg19_chr21$fullGenome, minoverlap = 1)

## Obtain fullCov object
fullCov <- list('chr21' = genomeDataRaw$coverage)
regionCov <- getRegionCoverage(fullCov=fullCov, regions=regions)
```

And now we can make the example plot as shown below.

```{r 'example_plot'}
## now make the plot
plotRegionCoverage(regions=regions, regionCoverage=regionCov, 
    groupInfo=genomeInfo$pop, nearestAnnotation=nearestAnnotation, 
    annotatedRegions=annotatedRegions, whichRegions=1:2,
    txdb = txdb_v31_hg19_chr21)
```



# JHPCE


You can also access the data locally using the function `local_metadata()` which works at [JHPCE](https://jhpce.jhu.edu/) or anywhere where you have re-created the files from this package. This returns a `data.frame()` which you can subset. It also inclused the R code for loading the data which you can do using  `eval(parse(text = local_metadata()$loadCode))` as shown below.

```{r 'local_metadata'}
## Get the local metadata
meta <- local_metadata()

## Subset to the data of interest, lets say hg19 TxDb for v31
interest <- subset(meta, RDataClass == 'TxDb' & Tags == 'Gencode:v31:hg19')

## Next you can load the data
if(file.exists(interest$RDataPath)) {
    ## This only works at JHPCE
    eval(parse(text = interest$loadCode))
    
    ## Explore the loaded object (would be gencode_v31_hg19_txdb in this case)
    gencode_v31_hg19_txdb
}
```

# Build objects

The objects provided by `GenomicState` through `r Biocpkg('AnnotationHub')` `r citep(bib[['AnnotationHub']])` were built using code like the one included below which is how the Gencode version 23 for hg19 files were built.

```{r 'build_local', eval = FALSE}
outdir <- 'gencode'
dir.create(outdir, showWarnings = FALSE)
 
## Build and save the TxDb object
gencode_v23_hg19_txdb <- gencode_txdb('23', 'hg19')
saveDb(gencode_v23_hg19_txdb,
    file = file.path(outdir, 'gencode_v23_hg19_txdb.sqlite'))

## Build and save the annotateTranscripts output
gencode_v23_hg19_annotated_genes<- gencode_annotated_genes(
    gencode_v23_hg19_txdb
)
save(gencode_v23_hg19_annotated_genes,
    file = file.path(outdir, 'gencode_v23_hg19_annotated_genes.rda'))

## Build and save the GenomicState
gencode_v23_hg19_GenomicState <- gencode_genomic_state(
    gencode_v23_hg19_txdb
)
save(gencode_v23_hg19_GenomicState,
    file = file.path(outdir, 'gencode_v23_hg19_GenomicState.rda'))
```

For more details check the source files:

```{r 'build_source'}
## R commands for building the files:
system.file('scripts', 'make-data_gencode_human.R',
    package = 'GenomicState')
## The above file was created by this one:
system.file('scripts', 'generate_make_data_gencode_human.R',
    package = 'GenomicState')
```

# Reproducibility

The `r Biocpkg('GenomicState')` package `r citep(bib[['GenomicState']])` was made possible thanks to:

* R `r citep(bib[['R']])`
* `r Biocpkg('BiocStyle')` `r citep(bib[['BiocStyle']])`
* `r Biocpkg('GenomeInfoDb')` `r citep(bib[['GenomeInfoDb']])`
* `r CRANpkg('knitcitations')` `r citep(bib[['knitcitations']])`
* `r CRANpkg('knitr')` `r citep(bib[['knitr']])`
* `r CRANpkg('rmarkdown')` `r citep(bib[['rmarkdown']])`
* `r Biocpkg('rtracklayer')` `r citep(bib[['rtracklayer']])`
* `r CRANpkg('sessioninfo')` `r citep(bib[['sessioninfo']])`
* `r CRANpkg('testthat')` `r citep(bib[['testthat']])`
* `r Biocpkg('GenomicFeatures')` `r citep(bib[['GenomicFeatures']])`
* `r Biocpkg('bumphunter')` `r citep(bib[['bumphunter']])`
* `r Biocpkg('derfinder')` `r citep(bib[['derfinder']])`
* `r Biocpkg('AnnotationDbi')` `r citep(bib[['AnnotationDbi']])`
* `r Biocpkg('IRanges')` `r citep(bib[['IRanges']])`
* `r Biocpkg('org.Hs.eg.db')` `r citep(bib[['org.Hs.eg.db']])`
* `r CRANpkg('glue')` `r citep(bib[['glue']])`
* `r Biocpkg('AnnotationHub')` `r citep(bib[['AnnotationHub']])`
* `r Biocpkg('AnnotationHubData')` `r citep(bib[['AnnotationHubData']])`
* `r Biocpkg('GenomicRanges')` `r citep(bib[['GenomicRanges']])`
    
Code for creating the vignette

```{r createVignette, eval=FALSE}
## Create the vignette
library('rmarkdown')
system.time(render('GenomicState.Rmd'))

## Extract the R code
library('knitr')
knit('GenomicState.Rmd', tangle = TRUE)
```


Date the vignette was generated.

```{r reproduce1, echo=FALSE}
## Date the vignette was generated
Sys.time()
```

Wallclock time spent generating the vignette.

```{r reproduce2, echo=FALSE}
## Processing time in seconds
totalTime <- diff(c(startTime, Sys.time()))
round(totalTime, digits=3)
```

`R` session information.

```{r reproduce3, echo=FALSE}
## Session info
library('sessioninfo')
options(width = 120)
session_info()
```

# Bibliography

This vignette was generated using `r Biocpkg('BiocStyle')` `r citep(bib[['BiocStyle']])`, `r CRANpkg('knitr')` `r citep(bib[['knitr']])` and `r CRANpkg('rmarkdown')` `r citep(bib[['rmarkdown']])` running behind the scenes.

Citations were made with `r CRANpkg('knitcitations')` `r citep(bib[['knitcitations']])`.

```{r vignetteBiblio, results = 'asis', echo = FALSE, warning = FALSE, message = FALSE}
## Print bibliography
bibliography()
```
